>

### 2019-04-15 paper/spinifex-paper.tex
pull the code usage into a new section, references pointing to the useage. grammerly check and knit.

While there are numerous methods that generate tour paths, this research
focuses on the manual tour. The manual tour was described in Cook and
Buja (\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user
to control the projection coefficients of a selected variable in a 2D
projection. The manipulation of these coefficients allows the analyst to
explore how sensitive the projection structure is to these
manipulations. As manual tours operate on only one variable at a time,
they particularly useful once a feature of interest has been identified.
One way to identify ``interesting'' features is with the use of a guided
tour (Hurley and Buja
\protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Guided tours
select a very specific path, that which approachs a projection that
optimizes an objective-function, similar to simulated annealing
(Kirkpatrick, Gelatt, and Vecchi
\protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
optimization of a function allows guided tours to rapidly identify
interesting projection features given the relatively large
parameter-space. After a projection of interest is identified an analyst
can then use the ``finer brush'' of the manual tour, by controlling the
contributions of individual variables to explore the sensitivity they
have to the projection structure.

This paper offers an implementation of manual tours with the package
\CRANpkg{spinifex}. This package also exends the graphics display of
\textbf{tourr} to the animation frameworks offered by \CRANpkg{plotly}
(Sievert \protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
\CRANpkg{gganimate} (Pedersen and Robinson
\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}). The algorithm
used to perform a radial manual tour is demonstrated on a toy data set.
Followed by an application and disscussion of the radial manual tour on
an aggregation of high energy physics experiments in a large
multivariate space.

The paper is organized as follows. Section \ref{sec:algorithm} explains
the algorithm using a toy dataset. Section \ref{sec:display} discusses
the display of the animation after the path has been generated. Section
\ref{sec:application} illustrates how this can be used for sensitivity
analysis applied to contemporary high energy physics. The last section,
\ref{sec:discussion} summarizes the work and discusses future research.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provided with a 2D projection, choose a variable to explore. This is
  called the ``manip'' variable.
\item
  Create a 3D manipulation space, where the manip variable has the full
  contribution.
\item
  Generate a rotation sequence which increases the norm of the
  coefficient to 1 and zeros it.
\end{enumerate}

The steps are described in more detail below with description,
equations, figures, and their code usage.

@@

The flea data, originally from Lubischew
(\protect\hyperlink{ref-lubischew_use_1962}{1962}), made available in
\textbf{tourr} is used to illustrate the algorithm. The data contains 74
observations across 6 variables, which are physical measurements of the
flea beetles. Each observation belong to one of three species.

@@

# Also see the vignette: 
vignette("spinifex_vignette")

f_data  <- tourr::rescale(flea[, 1:6])
f_path  <- save_history(f_data, guided_tour(holes()))
f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_cat   <- factor(flea$species)

@@

\begin{Schunk}
\begin{Sinput}
view_manip_space(basis = f_basis, 
                 manip_var = 5, 
                 labels = colnames(f_data))
\end{Sinput}
\end{Schunk}

@@

Post-multiply the manipulation space by the pre-defined rotation matrix
producing \textbf{RM}, the rotated manip space. This is one frame of
basis values, repeat this procress for each value in the sequences of
\(\theta\) and \(\phi\) for the complete animation.

@@

play_manual_tour(data = f_data,
                 basis = f_basis, 
                 manip_var = 5, 
                 angle = angle_speed,
                 col = f_cat,
                 pch = f_cat)
\end{Sinput}
\end{Schunk}

@@

\hypertarget{package-structure-and-functionality}{%
\section{Package structure and
functionality}\label{package-structure-and-functionality}}

@@

\hypertarget{walk-though-package-functions-in-terms-of-work-flow.}{%
\subsection{Walk though package functions in terms of work
flow.}\label{walk-though-package-functions-in-terms-of-work-flow.}}

@@

Tours, which are dynamic linear projection of multivariate data, play an
important role in data visualization; they extend the dimensionality of
visuals while data- and parameter-spaces become ever larger. This
research has modified the algorithm producing manual tours which and has
made this functionality available in package \textbf{spinifex}. The
package adds to \textbf{tourr}, extending the graphics offerings that
can be used to display tours.

>>>>

While there are numerous methods that generate tour paths, this research
focuses on the manual tour. The manual tour was described in Cook and
Buja (\protect\hyperlink{ref-cook_manual_1997}{1997}) and allows a user
to control the projection coefficients of a selected variable in a 2D
projection. The manipulation of these coefficients allows the analyst to
explore how sensitive the projection structure is to these
manipulations. As manual tours operate on only one variable at a time,
they particularly useful once a feature of interest has been identified.
One way to identify ``interesting'' features is with the use of a guided
tour (Hurley and Buja
\protect\hyperlink{ref-hurley_analyzing_1990}{1990}). Guided tours
select a very specific path, that which approaches a projection that
optimizes an objective-function, similar to simulated annealing
(Kirkpatrick, Gelatt, and Vecchi
\protect\hyperlink{ref-kirkpatrick_optimization_1983}{1983}). The direct
optimization of a function allows guided tours to rapidly identify
interesting projection features given the relatively large
parameter-space. After a projection of interest is identified an analyst
can then use the ``finer brush'' of the manual tour, by controlling the
contributions of individual variables to explore the sensitivity they
have to the projection structure.

This paper offers an implementation of manual tours with the package
\CRANpkg{spinifex}. This package also extends the graphics display of
\textbf{tourr} to the animation frameworks offered by \CRANpkg{plotly}
(Sievert \protect\hyperlink{ref-sievert_plotly_2018}{2018}) and
\CRANpkg{gganimate} (Pedersen and Robinson
\protect\hyperlink{ref-pedersen_gganimate:_2019}{2019}). The algorithm
used to perform a radial manual tour is demonstrated on a toy data set.
Followed by an application and discussion of the radial manual tour on
an aggregation of high energy physics experiments in a large
multivariate space.

The paper is organized as follows. Section \ref{sec:algorithm} explains
the algorithm using a toy dataset. Section \ref{sec:display} discusses
the display of the animation after the path has been generated. Section
\ref{sec:usage} explains the package functionality and code usage
following the order applied in the algorithm. Section
\ref{sec:application} illustrates how this can be used for sensitivity
analysis applied to contemporary high energy physics. The last section,
\ref{sec:discussion} summarizes the work and discusses future research.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provided with a 2D projection, choose a variable to manipulate. This
  is called the ``manip var''.
\item
  Create a 3D manipulation space, where the manip variable has the full
  contribution.
\item
  Generate a rotation sequence which increases the norm of the
  coefficient to 1 and zeros it.
\end{enumerate}

The steps are described in more detail below with descriptions,
equations, and figures are given inline. Function names are mentioned,
but code usage is discussed in more detail in section \ref{sec:usage}.

@@

The flea data, originally from Lubischew
(\protect\hyperlink{ref-lubischew_use_1962}{1962}), made available in
\textbf{tourr} is used to illustrate the algorithm. The data contains 74
observations across 6 variables, which are physical measurements of the
flea beetles. Each observation belongs to one of three species.

@@

@@

@@

Post-multiply the manipulation space by the pre-defined rotation matrix
producing \textbf{RM}, the rotated manip space. This is one frame of
basis values, repeat this process for each value in the sequences of
\(\theta\) and \(\phi\) for the complete animation.

@@

@@

\hypertarget{sec:usage}{%
\section{Package structure and functionality}\label{sec:usage}}

This section discusses the functions and code usage of the package
before delving into a domain-specific application of the manual tour.

@@

# Also see the vignette: 
vignette("spinifex_vignette")
\end{Sinput}
\end{Schunk}

\hypertarget{primary-package-functions}{%
\subsection{Primary package functions}\label{primary-package-functions}}

The primary functions that will aid users in performing their own manual
tours are broken into the following groupings: tour preparation, tour
animation, and utility.

\begin{Schunk}
\begin{table}[t]

\caption{\label{tab:functionsTable}Primary functions}
\centering
\begin{tabular}{lll}
\toprule
Class & Function & Description\\
\midrule
tour preparation & view\_basis & displays the reference frame of the basis\\
tour preparation & view\_manip\_space & displays the manipulation space\\
tour animation & play\_tour\_path & animates given tour path\\
tour animation & play\_manual\_tour & animates the manual tour algorithm\\
utility & render\_gganimate & uses gganimate to render the animation\\
\addlinespace
utility & render\_plotly & uses plotly to render the animation\\
\bottomrule
\end{tabular}
\end{table}

\end{Schunk}

In the following section describes and shows usage of the tour related
functions in parallel with the algorithm. The utility functions are
called by the tour animation functions, passed to the arguments
\texttt{render\_type}. Usage is demonstrated below and in the example
documentation.

\hypertarget{algorithm-code}{%
\subsection{Algorithm code}\label{algorithm-code}}

We'll start by initializing values including a standardized data set
(numeric columns only), a starting basis, a categorical variable for
point aesthetics, and a manip var. To get bearings on the projection,
start by observing the reference axes of the basis with
\texttt{view\_basis()}.

\begin{Schunk}
\begin{Sinput}
f_data  <- tourr::rescale(flea[, 1:6])                    ## standardize data
f_path  <- save_history(f_data, guided_tour(holes()))     ## produce guided tour
f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) ## end of guided tour
f_cat   <- factor(flea$species)                           ## categorical var

After becoming familiar with this space, select a manip var, the
variable to change the contributions of. Use
\texttt{view\_manip\_space()} View the new space with a dimension
orthogonal to the projection plane where the manip var has a full
contribution. This illustrates how the manip var is manipulated in a
high space not depicted in the manual tour.

\begin{Schunk}
\begin{Sinput}
f_mvar  <- 5  ## manip var numebr

view_manip_space(basis = f_basis, 
                 manip_var = f_mvar, 
                 labels = colnames(f_data))
\end{Sinput}
\end{Schunk}

Now we are ready to perform a manual tour on the selected variable. Use
\texttt{play\_manual\_tour()} to perform the algorithm as discussed
above, in section \ref{sec:algorithm}. The argument
\texttt{render\_type} can be changed to \texttt{render\_gganimate} for a
respective animation.

play_manual_tour(data = f_data,
                 basis = f_basis, 
                 manip_var = f_mvar, 
                 angle = angle_speed,
                 col = f_cat,
                 pch = f_cat)
\end{Sinput}
\end{Schunk}

Aside from the code used in the algorithm section a previously generated
tour path can also be animated via \texttt{play\_tour\_path()}. Utility
functions can also be passed into arguments: for example animation
format can be changed by setting
\texttt{render\_type\ =\ render\_gganimate}. View the guided tour path
as a gganimate function:

\begin{Schunk}
\begin{Sinput}
play_tour_path(tour_path = f_path,
               data = f_data,
               render_type = render_gganimate, 
               angle = angle_speed
)
\end{Sinput}
\end{Schunk}

@@

Tours, which are a dynamic linear projection of multivariate data, play
an important role in data visualization; they extend the dimensionality
of visuals while data- and parameter-spaces become ever larger. This
research has modified the algorithm producing manual tours which and has
made this functionality available in package \textbf{spinifex}. The
package adds to \textbf{tourr}, extending the graphics offerings that
can be used to display tours.

### 2019-04-15 paper/spinifex_paper.Rmd
pull the code usage into a new section, references pointing to the useage. grammerly check and knit.

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projection structure is to these manipulations. As manual tours operate on only one variable at a time, they particularly useful once a feature of interest has been identified. One way to identify "interesting" features is with the use of a guided tour [@hurley_analyzing_1990]. Guided tours select a very specific path, that which approachs a projection that optimizes an objective-function, similar to simulated annealing [@kirkpatrick_optimization_1983]. The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space. After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have to the projection structure.

<!-- Contributions -->
This paper offers an implementation of manual tours with the package
\CRANpkg{spinifex}. This package also exends the graphics display of **tourr** to the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019]. The algorithm used to perform a radial manual tour is demonstrated on a toy data set. Followed by an application and disscussion of the radial manual tour on an aggregation of high energy physics experiments in a large multivariate space.

<!-- Paper outline -->
The paper is organized as follows. Section \ref{sec:algorithm} explains the algorithm using a toy dataset. Section \ref{sec:display} discusses the display of the animation after the path has been generated. Section \ref{ref:usage} explains the though the code used in the algorithm section and discusses function structure for the rest of the package. Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics. The last section, \ref{sec:discussion} summarizes the work and discusses future research.

@@

1. Provided with a 2D projection, choose a variable to explore. This is called the "manip" variable.
2. Create a 3D manipulation space, where the manip variable has the full contribution.
3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.

The steps are described in more detail below with description, equations, and figures. Function names are mentioned inline, but the code is aggregated in

@@

<!-- describe data. -->
The flea data, originally from @lubischew_use_1962, made available in **tourr**  is used to illustrate the algorithm. The data contains 74 observations across 6 variables, which are physical measurements of the flea beetles. Each observation belong to one of three species.

@@

Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space. This is one frame of basis values, repeat this procress for each value in the sequences of $\theta$ and $\phi$ for the complete animation.

@@

@@

## Algorithm

We'll start by initializing values including: standardizing our data set (numeric coulmns only), provide a basis, a categorical variable for point asthetics, and specifiy the manip var. Then we the references axes of the basis with `view_basis()`.

@@

After becoming familar with the space, select a

@@

@@

## Walk though package functions in terms of work flow.

@@

<!-- summary and segue  -->
Tours, which are dynamic linear projection of multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger. This research has modified the algorithm producing manual tours which and has made this functionality available in package **spinifex**. The package adds to **tourr**, extending the graphics offerings that can be used to display tours.

>>>>

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projection structure is to these manipulations. As manual tours operate on only one variable at a time, they particularly useful once a feature of interest has been identified. One way to identify "interesting" features is with the use of a guided tour [@hurley_analyzing_1990]. Guided tours select a very specific path, that which approaches a projection that optimizes an objective-function, similar to simulated annealing [@kirkpatrick_optimization_1983]. The direct optimization of a function allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space. After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables to explore the sensitivity they have to the projection structure.

<!-- Contributions -->
This paper offers an implementation of manual tours with the package
\CRANpkg{spinifex}. This package also extends the graphics display of **tourr** to the animation frameworks offered by \CRANpkg{plotly} [@sievert_plotly_2018] and \CRANpkg{gganimate} [@pedersen_gganimate:_2019]. The algorithm used to perform a radial manual tour is demonstrated on a toy data set. Followed by an application and discussion of the radial manual tour on an aggregation of high energy physics experiments in a large multivariate space.

<!-- Paper outline -->
The paper is organized as follows. Section \ref{sec:algorithm} explains the algorithm using a toy dataset. Section \ref{sec:display} discusses the display of the animation after the path has been generated. Section \ref{sec:usage} explains the package functionality and code usage following the order applied in the algorithm. Section \ref{sec:application} illustrates how this can be used for sensitivity analysis applied to contemporary high energy physics. The last section, \ref{sec:discussion} summarizes the work and discusses future research.

@@

1. Provided with a 2D projection, choose a variable to manipulate. This is called the "manip var".
2. Create a 3D manipulation space, where the manip variable has the full contribution.
3. Generate a rotation sequence which increases the norm of the coefficient to 1 and zeros it.

The steps are described in more detail below with descriptions, equations, and figures are given inline. Function names are mentioned, but code usage is discussed in more detail in section \ref{sec:usage}.

@@

<!-- describe data. -->
The flea data, originally from @lubischew_use_1962, made available in **tourr**  is used to illustrate the algorithm. The data contains 74 observations across 6 variables, which are physical measurements of the flea beetles. Each observation belongs to one of three species.

@@

Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space. This is one frame of basis values, repeat this process for each value in the sequences of $\theta$ and $\phi$ for the complete animation.

@@

This section discusses the functions and code usage of the package before delving into a domain-specific application of the manual tour.

@@

## Primary package functions 
<!-- in terms of work flow. -->

The primary functions that will aid users in performing their own manual tours are broken into the following groupings: tour preparation, tour animation, and utility.

```{r functionsTable, echo=F}
library(kableExtra)

funcs <- rbind(
  c("tour preparation", "view_basis", "displays the reference frame of the basis"),
  c("tour preparation", "view_manip_space", "displays the manipulation space"),
  c("tour animation", "play_tour_path", "animates given tour path"),
  c("tour animation", "play_manual_tour", "animates the manual tour algorithm"),
  # c("utility", "col_of", "maps a variable to the point color"),
  # c("utility", "pch_of", "maps a variable to the point character (shape)"),
  c("utility", "render_gganimate", "uses gganimate to render the animation"),
  c("utility", "render_plotly", "uses plotly to render the animation")
)
colnames(funcs) <- c("Class", "Function", "Description")

kable(funcs, "latex", caption = "Primary functions", booktabs = T)
```

In the following section describes and shows usage of the tour related functions in parallel with the algorithm. The utility functions are called by the tour animation functions, passed to the arguments `render_type`. Usage is demonstrated below and in the example documentation.

## Algorithm code

We'll start by initializing values including a standardized data set (numeric columns only), a starting basis, a categorical variable for point aesthetics, and a manip var. To get bearings on the projection, start by observing the reference axes of the basis with `view_basis()`.

@@

After becoming familiar with this space, select a manip var, the variable to change the contributions of. Use `view_manip_space()` View the new space with a dimension orthogonal to the projection plane where the manip var has a full contribution. This illustrates how the manip var is manipulated in a high space not depicted in the manual tour.

@@

Now we are ready to perform a manual tour on the selected variable. Use `play_manual_tour()` to perform the algorithm as discussed above, in section  \ref{sec:algorithm}. The argument `render_type` can be changed to `render_gganimate` for a respective animation.

@@

Aside from the code used in the algorithm section a previously generated tour path can also be animated via `play_tour_path()`. Utility functions can also be passed into arguments: for example animation format can be changed by setting `render_type = render_gganimate`. View the guided tour path as a gganimate function:

```{r guidedTourCODEEXAMPLE, eval=F, echo=T}
play_tour_path(tour_path = f_path,
               data = f_data,
               render_type = render_gganimate, 
               angle = angle_speed
)
```

@@

<!-- summary and segue  -->
Tours, which are a dynamic linear projection of multivariate data, play an important role in data visualization; they extend the dimensionality of visuals while data- and parameter-spaces become ever larger. This research has modified the algorithm producing manual tours which and has made this functionality available in package **spinifex**. The package adds to **tourr**, extending the graphics offerings that can be used to display tours.
